---
title: "Using the smhiAPI for fetching and plotting data"
author: "Anton Persson & Emil Klasson Svensson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{smhiAPI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

smhiAPI is a package created during a R-course at Linkoping University. The package is supposed to collect data from SMHI for the four last months via XML, and has functions both to return a dataframe with the requested data, and to plot the data versus time.

The users are supposed to be able to collect data from every weather station that collects data continously in Sweden. The users can also specify what kind of data they want, for example air temperature and air humidity. 

## Vignette Info

The smhiAPI package includes four functions. The functions get\_parameters and get\_station has no input argument. The functions smhi\_data and smhi\_plot has two arguments, "station" and "parameter". 

get\_station is just supposed to show the user what stations with fresh data are available at the moment, and the stations ID respectively. get\_parameter shows a list of the available parameters, and their respective ID. 
I.E, they are used to look up what the input in the smhi\_data and smhi\_plot should be.


## Examples of how to use the smhiAPI package

If you don't have specific knowledge about the data from SMHI, you probably won't know what ID belongs to what weather station. This ID is needed to be able to collect data from SMHI. That's why it's a good start to call the get\_station, to find the ID of the weather station you want to look into to.


```{r, echo=TRUE, message = FALSE, warning = FALSE }
library(smhiAPI)
head(get_station()) #show the first six stations

```

Since we in this example assume that you want to dig into the air temperature in Arvidsjaur, we can see that the ID for the Arvidsjaur station is 159880. That's our station argument. 
Now we wanna see if the air temperature is available as the parameter.

```{r, echo=TRUE, message = FALSE, warning = FALSE}
get_parameters()
```

The air temperature is available as a parameter! Let's assume that we want the air temperature as a average daily value. Then we can see that the parameter we're searching for is 2. So now we get our input, station=159880 and parameter=2. Let's try to collect some data!

```{r, echo=TRUE, message = FALSE, warning = FALSE}
mydata<-smhi_data(station=159880,parameter=2)

```

So, now we got our data. mydata is a dataframe that consists the average daily air temparature and the respective date for each day the last four months. Now, we want to plot it to see how the the temperature has developed over the last four months. It's time to use the smhi\_plot function!

```{r, echo=TRUE, message = FALSE, warning = FALSE, fig.height=7, fig.width=7}
smhi_plot(station=159880,parameter=2)
```

So, we can see that the autumn of 2016 has arrived to Arvidsjaur, when this vignette was written.


##Contact info

If you find any problems with the package or it's documentation, please send us some feedback or critique. Contact Anton Persson, antpe404@student.liu.se, or Emil Klasson Svensson, emisv463@student.liu.se.